### 2018-06-16 CS761-RandomizedAlgorithms/assignments/a2.tex
writing 3

\lhead{\today}
\rhead{W. Justin Toth CS761-Randomized Algorithms Assignment n2} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

@@

%Problem 4
\newpage
\section{}

>>>>

\lhead{\today}
\rhead{W. Justin Toth CS761-Randomized Algorithms Assignment 2} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

@@

Suppose we have a list $L$ of $n$ distinct numbers, whose sorted list would we $\{x_1,\dots, x_n\}$. We want to find a $k$-approximate median. I.e. we want to return some $x \in [x_{\frac{n}{2} - k}, x_{\frac{n}{2} + k}]\cap L$. Consider sampling $m$ numbers from $L$ and returning the median $c$ of these $m$ numbers as the true median of $L$.
Let $L'$ be list of $m$ sampled numbers. For each $i \in L'$ let $X_{i,<}$ be an indicator variable for the event that $i \in \{x_1, \dots, x_{\frac{n}{2} -k -1}\}$. Then, since we sample without replacement,
$$Pr(X_{i, <} =1) = (\frac{n}{2} -k -1) / n = \frac{1}{2} - \frac{k+1}{n}$$
Let $X_<$ be the number of elements in $L'$ which lie in $\{x_1, \dots, x_{\frac{n}{2} -k -1}\}$. Then $X_< = \sum_{i \in L'} X_{i,<}$.
Let $X_>$ be the number of elements in $L'$ which lie in $\{x_{\frac{n}{2} + k + 1}, \dots, x_n\}$. Since we sample uniformly and
	$$|\{x_{\frac{n}{2} + k + 1}e, \dots, x_n\}| =  |\{x_1, \dots, x_{\frac{n}{2} -k -1}\}| = \frac{n}{2} - k - 1$$
we observe that $X_<$ and $X_>$ have the same distribution.
Now we see that $c$ fails to be a $k$-approximate median for $L$ if and only if $X_< \geq \frac{m}{2}$ or $X_> \geq \frac{m}{2}$. Since $X_<$ and $X_>$ have the same distribution, the probability of failure is thus
$$2Pr(X_< \geq \frac{m}{2}).$$
We plan to bound this using a Chernoff bound. First we compute the expectation,
$$E[X_<] = \sum_{i\in L'} Pr(X_{i,<} = 1)  = \sum_{i \in L'} (\frac{1}{2} - \frac{k+1}{n}) = \frac{m(n-2(k+1))}{2n}.$$
So 
$$Pr(X_< \geq \frac{m}{n}) = Pr(X_< \geq (1+\delta) E[X_<])$$
where $$\delta = \frac{2}{\frac{n-2(k+1)}{2}}-1=\frac{4-(n-2(k+1))}{n-2(k+1)}$$
%Problem 4
